% Created 2023-12-29 Fri 21:11
% Intended LaTeX compiler: pdflatex
\documentclass[a4paper]{article}

\usepackage{../preamble}
\author{Nicky}
\date{\today}
\title{Making an Exercise and Solution Environment}
\begin{document}

\maketitle
\section{The Problem}
\label{sec:org13ad8cf}

In a number of my org mode files I include exercise with hints and solutions.
In the pdf files, which I make via \LaTeX, I use the \texttt{answers} package print the exercises in the main text and the hints and solutions at the end of the file.
When exporting the file to \texttt{html} via my homepage I would like to see the exercises, hints and solutions printed like so:

\begin{exercise}
How to make an excercise environment like this?
\begin{hint}
And a hint like this?
\end{hint}
\begin{solution}
To see how to do this, read on.
\end{solution}
\end{exercise}
\section{My Environment}
\label{sec:org1a6fb2f}

I use \texttt{emacs} and \texttt{org mode} to write files, and \texttt{nikola} to make my homepage.
After some searching on the web, I found the \href{http://alhassy.com/org-special-block-extras/}{\texttt{org-special-blocks-extra}} package which proved promissing for my goal.
Here are the steps I followed.

First I installed the \texttt{org mode} plugin for \texttt{nikola}. The \texttt{orgmode.py} file shows that the next command can used to convert an \texttt{org} to an \texttt{html} file that \texttt{nikola} can read.
\begin{minted}[]{shell}
/usr/bin/emacs --batch -l init.el --eval '(nikola-html-export "input.org" "output.html")'
\end{minted}
The org mode plugin also provides an \texttt{init.el} file that I used as a starting point.

\begin{exercise}
The \texttt{init.el} file contains the line \texttt{(setq package-load-list '((htmlize t)))}. This should be removed. Why?
\begin{solution}
I need to install \texttt{org-special-blocks-extra}, and this command prevents that.
\end{solution}
\end{exercise}

Since I need functionality from the special org blocks, I have to update the \texttt{init.el} anyway, so here it is.
\section{The adapted \texttt{init.el}}
\label{sec:org4c61ab0}

Since I use  \texttt{straight} and \texttt{use-package}, I need to put this on the top of the file.

\begin{minted}[]{common-lisp}
(defvar bootstrap-version)
(let ((bootstrap-file
       (expand-file-name "straight/repos/straight.el/bootstrap.el" user-emacs-directory))
      (bootstrap-version 5))
  (unless (file-exists-p bootstrap-file)
    (with-current-buffer
        (url-retrieve-synchronously
         "https://raw.githubusercontent.com/raxod502/straight.el/develop/install.el"
         'silent 'inhibit-cookies)
      (goto-char (point-max))
      (eval-print-last-sexp)))
  (load bootstrap-file nil 'nomessage))
(setq package-enable-at-startup nil)

(straight-use-package 'use-package)
(setq straight-use-package-by-default t)
\end{minted}

Next I load two packages.

\begin{minted}[]{common-lisp}
(use-package htmlize
  :ensure t)

(use-package org-special-block-extras
  :ensure t
  ;; All relevant Lisp functions are prefixed ‘o-’; e.g., `o-docs-insert'.
  :custom
    (o-docs-libraries
     '("~/org-special-block-extras/documentation.org")
     "The places where I keep my ‘#+documentation’"))
\end{minted}
\end{document}
